---
title: "Penguins ANOVA"
author: "Ben Edwards"
date: "11/6/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Load packages and read in penguins.csv data

```{r, include = FALSE}

library(tidyverse)
library(car)
library(onewaytests)

penguins <- read_csv("penguins.csv")

```

Our Question:

Is there sif dif in mean difference in mean flipper length for female pengins by species (Chinstrap, Gentoo, and Adelie)?

2. Do some exploring

New funcitons: with + table for easy counting over multple factors

update species name using case_when

```{r}

counts <- with(penguins, table(species, sex))

penguins2 <- penguins %>%
  mutate(common_name = 
           case_when(
             species == "Chinstrap penguin (Pygoscelis antarctica)" ~ "Chinstrap",
             species == "Gentoo penguin (Pygoscelis papua)" ~ "Gentoo",
             species == "Adelie Penguin (Pygoscelis adeliae)" ~ "Adelie")) %>%
select(-species) %>%
filter(sex == "FEMALE")
```

-Exploratory graphs
-Summary statistics
-Levene's tests for equal variances
- one-way ANNOVA
- post-hoc testing

```{r}

hists <- ggplot(penguins2, aes(x = flipper_length)) + 
  geom_histogram(aes(fill = common_name)) + 
  facet_wrap(~ common_name)

hists

qq <- ggplot(penguins2, aes(sample = flipper_length)) +
  geom_qq() +
  facet_wrap(~ common_name)

qq


# Levene's test for equal variances (> 2 groups....if you have two groups, use the f-test for equal variances)
# HO: Variances are equal
# HA: Variances are not eual

# We'll se the function leaveneTest from teh 'car' package

penguin_levene <- leveneTest(flipper_length ~ common_name, data = penguins2)
penguin_levene

# We retain the null hypothesis of equal variances (p= 0.06)

variances <- penguins2 %>%
  group_by(common_name) %>%
  summarize(
    variance = var(flipper_length))

variances

```

#3. One-way ANOVA

Single factor: common_name (species)
Number of levels in the factor: 3 (Chinstrap, Gentoo, Adelie)
Random Variable: flipper_length (mm)

Is there a sig diff in mean flipper length (mm) for female Chinstrap, Gentoo and Adelie penguins?

```{r}
# HO: There is no difference in the means of flipper lengths accros all species
# HA: At least 2 means differ significantly (mean flipper length)

penguin_aov <- aov(flipper_length ~ common_name, data = penguins2)
summary(penguin_aov)

# At least two samples were taken from populations with different means. Which ones are different? All three are different from each other? Or somethign else?

# post-hoc testing using tukey's hst

penguin_ph <- TukeyHSD(penguin_aov)
penguin_ph

```
#Mean flipper lengths for Adelie, Gentoo, and Chinstrap penguins are sig diff (F(df) = statistic, p < )
#This means if we take samples from these 3 groups and they actually come from samples with the same mean flipper length there is almost no chance that this happens by random chance, so we conclude that the populations probably actually have different means. AT LEAST 2 SAMPLES WERE TAKEN FROM POPULATIONS WITH DIFFERENT MEANS.
##There is a sig diff in flipper length between the 3 penguin species.

#Mean flipper lengths are all diff

